---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# DIGCLASS

<!-- badges: start -->
<!-- badges: end -->

The `DIGCLASS` R package aims to make translation between occupational social classes easy and comprehensive. It facilitates the translation of the International Standard Classification of Occupations (ISCO) from 1968, 1988 and 2008 to a wide range of social class schemes.

This package is a work in progress and has implemented currently these translations:

* **ISCO68**
  - [X] ISCO68 to ISCO88
  - [X] ISCO68 to ISCO08 # If this one is done then ISCO08 -> ISCO68 should be done
  - [X] ISCO68 to ISEI
  - [X] ISCO68 to SIOPS
  - [X] ISCO68 to EGP
  - [X] ISCO68 to EGP11

* **ISCO88**
  - [X] ISCO88 to ISEI
  - [X] ISCO88 to SIOPS
  - [X] ISCO88 to MPS88
  - [X] ISCO88 to EGP
  - [X] ISCO88 to EGP11
  - [X] ISCO88 to OESCH16
  - [X] ISCO88 to ISCO88COM
  - [X] ISCO88 to ISCO08
  - [X] ISCO88 to ISCO68

* **ISCO08**
  - [X] ISCO08 to ISEI
  - [X] ISCO08 to SIOPS
  - [X] ISCO08 to OESCH16
  - [ ] ISCO08 to ISCO68 (not available)
  - [X] ISCO08 to ISCO88



## Installation

You can install the development version of DIGCLASS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cimentadaj/socialclasses")
```

## Example

Here's an example of translating ISCO68 to ISEI and EGP class schemas:

```{r,message=FALSE, warning=FALSE }
library(DIGCLASS)
library(dplyr)

# Internal data for the European Social Survey round 5
# containing different ISCO variables
ess %>%
  mutate(
    isei = isco68_to_isei(isco68),
    egp = isco68_to_egp(isco68, self_employed, emplno),
    egp_labels = isco68_to_egp(isco68, self_employed, emplno, label = TRUE)
  )
```


## Steps to add a new translation

1. Add two csv files respectively in `data-raw/social_classes/labels/` and `data-raw/social_classes/translation/` containing the labels and translation for the two schemas.

2. Run the script `data-raw/social_classes.R` (with the root directory in `data-raw/`)

3. Add a new function inside `R/` with the convention `{origin}_to_{destination}()` where origin and destination are the class schemas we're translating. Please have a look at other translation to recycle common functions to do translations.

4. Add proper documentation to the function


## Other R packages

This package has benefitted greatly from other open source packages that already pave the way for translation between social class schemas. In particular, we've learned a lot and borrowed code from all of these packages:

- [ISCOGEN](https://github.com/benjann/iscogen): Stata package
- [SocialPosition](https://cran.r-project.org/web/packages/SocialPosition/index.html): R package
- [occupar](https://github.com/DiogoFerrari/occupar/): R package

## TODO:
- Oesch schemas have `.` inside the classes, what to do with those? We get NAs introduced when converting to numeric.
- Oesch seems to have other labels which are shorter but no requivalence
- From parsed txt file only ESEC missing
- Need to make sure which ESEC you parsed from txt matches the ones in their doc. They want more ESEC so you need to map them and add additional ESEC.
- Add recoding of major/minor codes for all ISCO.
- Adapt E.O wright schema
- Create vignette examples with already existing codifications
- Plan meeting with them to show current work
- Improve docs on each translation maybe mentioning what each  class schemas is and pointing to the source and the website.

