---
title: "Case study: Exploring translations between social classes using the ESS data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{case-study-translations-social-classes-ess.Rmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Exploring translations between social classes

Let's suppose you've received a dataset which contains the International Standard Classification of Occupations (ISCO) on occupations of respondents in a survey. In particular, you've got only the ISCO classification for the year 1968, which is a bit outdated. With the package `DIGCLASS` you can translate that classiciation to many others. Let's load our package and use the data `ess` which contains ISCO68 from the European Social Survey.


```{r setup}
library(DIGCLASS)
library(dplyr)

ess
```

As you can see, this dataset contains `isco68`, `isco88` and `isco08`. Assume for a second that you **do not** have `isco88` and `isco08` and you'd like to translate `isco68` to other class schemas. In `DIGCLASS` you will find all functions that translate `isco68` begin with `isco68_`. So if you typed on your text editor `isco68_` and hit `TAB`, presumably you'll get auto-completion for all possible translations. Here's a selected number of translations you can get for `isco68`:

```{r}
# Set `labels` to `FALSE` to obtain
# the translated number instead of the label
labs <- FALSE

# For some class schemas we use the number of employees that
# the respondent has in their job. Here we're just recoding
# some values that are missing.
ess <- ess %>% mutate(emplno = if_else(emplno > 10000, 0, emplno))


# Translate `isco68` to many class schemas
ess %>%
  select(isco68, self_employed, emplno) %>%
  transmute(
    isco68,
    isco88 = isco68_to_isco88(isco68, label = labs),
    isco08 = isco68_to_isco08(isco68, label = labs),
    siops = isco68_to_siops(isco68),
    isei = isco68_to_isei(isco68),
    egp = isco68_to_egp(isco68, self_employed, emplno, label = labs),
  )
```


As you can see, we got all translations done rather quickly.

`DIGCLASS` aims to have a consistent interface, meaning that if you happen to have ISCO88 instead of ISCO68, the same convention for all functions is used:

```{r}
ess %>%
  select(isco88, self_employed, emplno) %>%
  transmute(
    isco88,
    isco68 = isco88_to_isco68(isco88, label = labs),
    isco88com = isco88_to_isco88com(isco88, label = labs),
    isco08 = isco88_to_isco08(isco88, label = labs),
    siops = isco88_to_siops(isco88),
    isei = isco88_to_isei(isco88),
    mps = isco88_to_mps(isco88),
    egp = isco88_to_egp(isco88, self_employed, emplno, label = labs),
    oesch = isco88_to_oesch(isco88, self_employed, emplno, label = labs)
  )
```

Similarly, you can apply transformations for ISCO08 following the same skeleton. For this one let's see `labs = TRUE` to check the labels:

```{r}
labs <- TRUE

ess %>%
  select(isco08, self_employed, emplno) %>%
  transmute(
    isco08,
    isco88 = isco08_to_isco88(isco08, label = labs),
    siops = isco08_to_siops(isco08),
    isei = isco08_to_isei(isco08),
    oesch = isco08_to_oesch(isco08, self_employed, emplno, label = labs)
  )

```

Notice that some classes translate to the labels (`isco88` and `oesch`) but others like `isei` and `siops` are actual numeric translations. That's because they don't have a `label` argument; they are continuous scales.

**Note that swapping the first part of the function (`isco88_` or `isco68_`) for another ISCO will not always work because not all `isco**` have the same translations. Be sure to check the reference of functions to see which translations are available for each ISCO class schema.**
