---
title: "Repairing an ISCO variable"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{repairing_isco_input}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Short Summary

When reading data into R, ISCO variables tend to loose information when read as a number. Be sure to specify that ISCO variables be read as character vectors instead of numeric ones. You can do that for CSV files and Stata formats like this:

```{r, eval = FALSE}
library(readr)
# For CSV, specify `col_types`
isco_ready <- read_csv(isco_csv, col_types = list(ISCO08 = col_character(), ISCO88 = col_character()))

# For Stata
read_dta("ess.dta", col_types = c(ISCO08 = "character"))
```

## Repairing ISCO variables

ISCO variables can become corrupted when being read into R. Let me show you a typical example. Suppose we have a CSV file with data on ISCO occupations:


```{r setup}
library(DIGCLASS)
library(readr)

isco_csv <- "ISCO08,ISCO88
0110,0110
0210,0110
0310,0110
0100,0110
0200,0110
0300,0110
1000,1000
1100,1100
1110,1100
1111,1110"
```

This CSV data is in a string in R and we need to read it with `read_csv`. It is quite common for these ISCO columns to be read as a numeric column when using functions such as `read_csv` or `read_dta` (for Stata files). Let's see what happens when we read a column as a numeric:

```{r}
isco_ready <- read_csv(isco_csv, col_types = list(.default = col_double()))
```

The first 6 rows lost the first `0`. This is quite dangerous because right now we don't know if the occupation `110` is actually `0110` or `1100`. Since we can't check every single occupation in our dataset to confirm that it kept all the information of each digit, `DIGCLASS` has the function `repair_isco` which tells you whether your ISCO column is ready for analysis. Let's run it:

```{r, error = TRUE}
repair_isco(isco_ready$ISCO08)
```

The functions tells us precisely that this ISCO variable has 3 and 4 digit occupations and that it needs to be normalized. We can fix that be reading the data specifying the columns to be as a character:

```{r}
isco_ready <- read_csv(isco_csv, col_types = list(ISCO08 = col_character(), ISCO88 = col_character()))
isco_ready$ISCO08 <- repair_isco(isco_ready$ISCO08)
```

Similarly for Stata files, you should always specify that the ISCO variable should be read as a character like this:

```{r, eval = FALSE}
read_dta("ess.dta", col_types = c(ISCO08 = "character"))
```

