---
title: "Translating between hierarchies for ISCO68, ISCO88 and ISCO08"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{translating_between_hierarchies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Translating between hierarchies

Most surveys that contain occupation related variables have 4 digit ISCO occupations. What does that mean? That you're working with the most fine-grained definition of an occupation. In some cases, you want to work with aggregated groups. Instead of knowing something about a mathematician, you'd rather group all math related occupation into a "Scientist" category. `DIGCLASS` has this implemented following the rules of each ISCO schema. Let's load `DIGCLASS`:

```{r setup}
library(DIGCLASS)
library(dplyr)
```

In ISCO parlance, the most granular occupation have what it's called 4 digits. This means that the occupation 4 non-zero digits. Occupation 2111 is a 4 digit occupation because it does not contain any zeroes. In contrast, 2110 is the "parent" category of 2111. To make it even more simple, think that 2111 is the occupation "Physicists and astronomers" while 2110 is "Physicists, chemists and related professionals". You can intuitively group physicists in that broader category. Similarly, the occupation 2110 or "Physicists, chemists and related professionals" is nested within the more broader group 2100 or "Physical, mathematical and engineering science professionals". Finally, the broadest group is 2000, for which the general group definition is "Professionals.

This was just an intuitive explanation of how ISCO codes works. You don't have to remember what each category is. You can always look up these values yourself for better understanding but `DIGCLASS` will do the work of translating everything for you. The important thing to remember is that something you'll want to group fine grained occupations into more broader occupation categories. An example would be that all categories that are within "Physicists, chemists and related professionals" are grouped together. This means that we "convert" the column from 4 digits into 2 digits, for example. In `DIGCLASS` you can do that with the function `isco*_swap` where `*` is the ISCO of preference. Let's look at the ISCO variables we have in the ESS data in `DIGCLASS`:

```{r}
ess
```

All three ISCO variables are in four digits but we can convert them to three digits:

```{r}
ess %>%
  transmute(
    isco88,
    isco88_three = isco88_swap(isco88, from = 4, to = 3)
  )
```

As you can see, the three digit translation always has a zero, meaning that it was translated into a broder group. We can do the same for an even broader group, translating from 4 to 2 digits:

```{r}
ess %>%
  transmute(
    isco08,
    isco08_two = isco08_swap(isco08, from = 4, to = 2)
  )
```

We can see that the two digit translation is a broader category than the original four digit occupation. Note that we can translate everything from 4 to 1 but not the other way around:

```{r, error = TRUE}
ess %>%
  transmute(
    isco08,
    isco08_two = isco08_swap(isco08, from = 2, to = 4)
  )
```

That's because we can't translate a more broader group into a finer occupation because it could be many specific occupation within a broder group. Finally, do note that for ISCO68, there are some 1 digit groups missing (0000 and 1000 don't have a broader category), so when you translate from any digit to the 1 digit in ISCO68 you might some missing values for occupation within the major group 0000 and 1000:

```{r}
ess %>%
  transmute(
    isco68,
    isco68_one = isco68_swap(isco68, from = 4, to = 1)
  )
```

Note that the 1 digit groups `2000`, `3000`, `5000` and `8000` are translated correctly. Yet the 1 digit group `1000` or `0000` are never translated because they don't exist in ISCO68. `DIGCLASS` makes the translation either way but note that you'll lose that information when you translate it to other schemas because it's an `NA`.

## Using translated hierarchies for translation between schemas

`isco*_swap` are important functions because some translations require ISCO variables to be in different digits. For example, to translate ISCO08 to the ESEC class schema, ISCO08 needs to be in 3-digits. How would that translation look like? Here's an example:

```{r}
library(dplyr)

# convert isco08 to three digits
ess$isco08_three <- isco08_swap(ess$isco08, from = 4, to = 3)

ess %>%
  transmute(
    isco08_three,
    esec = isco08_to_esec(
      isco08_three,
      is_supervisor,
      self_employed,
      emplno,
      label = FALSE
    )
  )
```

Similarly, ESEC has another translation but based on ISCO08 being 2-digits. Here's an example:

```{r}
# convert to two digits
ess$isco08_two <- isco08_swap(ess$isco08, from = 4, to = 2)

ess %>%
  transmute(
    isco08_two,
    esec = isco08_two_to_esec(
      isco08_two,
      is_supervisor,
      self_employed,
      emplno,
      label = FALSE
    )
  )
```


As you can see, `isco*_swap` are functions that serve and facilitate a common task in ISCO translations.

## Filling ISCO codes with trailing zeros

When working with ISCO codes at different digit levels, you may need to convert them to a specific digit format for certain functions. The `isco_fill()` function makes this easy by automatically appending the appropriate number of trailing zeros:

```{r}
# Example with 2-digit ISCO codes filled to 4-digit (default)
isco_2digit <- c("11", "21", "31", "61")
isco_filled <- isco_fill(isco_2digit)
print(isco_filled)

# Now you can use with OEP calculation
oep_scores <- isco08_to_oep(isco_filled)
print(oep_scores)
```

You can also fill to other digit levels:

```{r}
# You can also fill to other digit levels
isco_1digit <- c("1", "2", "6")
isco_fill(isco_1digit, digits = 3)  # Fill to 3-digit

isco_fill(isco_1digit, digits = 2)  # Fill to 2-digit
```

This is particularly useful when your data contains ISCO codes at the 1, 2, or 3-digit level but you need to use functions that expect 4-digit codes, such as `isco*_to_oep()`. The function follows the same logic as CROSSWALK/ISCOGEN in Stata.

Here's a typical workflow combining `isco_fill()` with other DIGCLASS functions:

```{r}
# Simulate some 2-digit ISCO data
isco_data <- data.frame(
  respondent_id = 1:4,
  isco08_2digit = c("11", "21", "31", "61"),
  self_employed = c(0, 0, 1, 0),
  emplno = c(0, 0, 3, 0)
)

# Fill to 4-digit and calculate OEP and OESCH
isco_data %>%
  mutate(
    isco08_4digit = isco_fill(isco08_2digit),
    oep = isco08_to_oep(isco08_4digit),
    oesch = isco08_to_oesch(isco08_4digit, self_employed, emplno, label = TRUE)
  )
```

The `isco_fill()` function is designed to work seamlessly with the rest of the DIGCLASS ecosystem, making it easy to prepare your ISCO data for analysis regardless of the original digit level.
